Section Rings of $\bq$-divisors on minimal rational surfaces

### 2017-11-24 surfaces-write-up/MRL/MRL-surfaces-responses-to-referee-3.tex

\def\listtodoname{List of Todos}
\def\listoftodos{\@starttoc{tdo}\listtodoname}

\title{Summary of Changes in response to Referee's Report}
\author{Aaron Landesman, Peter Ruhm, and Robin Zhang}

\begin{document}

\maketitle
\section{Changes Made}
Here are some responses to the referee's suggestions.
Item's below are numbered to match the referee report.
We applied nearly all changes suggested by the referee, though most need no
comment. Here, we make some brief comments on how we applied certain suggested
changes.
\begin{enumerate}
	\item[(2)] 
We have now added in a detailed example 
\autoref{example:effective-example}
explaining how to apply our method of proof to actually work out
the generators and relations for a section ring
associated to an effective divisor on $\bp^m$.
\item[(10)]
Response to the concern about base change along inseparable extensions:
We added Remark \autoref{rem:algebraically-closed}
explaining why purely inseparable extensions pose
no addition issue. The point is that cohomology commutes with flat base change, and extensions of fields
are always flat, so the dimensions of the graded pieces of the ring will be preserved. Therefore generators
and relations are preserved under arbitrary base field extension. Of course,
canonical divisors of root stacks may not behave well with respect to base change (as we learned from a version of Voight--Zureick-Brown's forthcoming book available on Voight's website)
which is why inseparable extensions
may cause problems for canonical rings.
But, for the purposes of this article, we fix a divisor on a scheme over some base field, and the structure of that particular section
ring is unchanged upon base change to the algebraic closure.

\item[(14) and (18)]
We explained in both the projective space case and Hirzebruch case that this assumption was nonessential, adding
\autoref{cor:proj-generators-relations} for the projective case
and \autoref{rem:hirz-generators-relations-hypothesis} for the Hirzebruch case.
You are correct that we could have easily changed coordinates and reordered the $f_i$ so that the first $m+1$ $f_i$ were $x_0, \ldots, x_m$, but
it seems we may as well state it with arbitrary independent polynomials, so as not to require such a change of basis.
\item[(22)]
	We moved the notation to its own section after the introduction,
	as most of the notation was not necessary to state
	our main theorems.
\item[(24)]
	You're right that the proof of \autoref{prop:pm-span-and-basis}
		was poorly phrased, and we have tried to clarify this now.
The point is that we are only varying $f_0, \ldots, f_m,$ so
the resulting monomials in the $f_i$ will be independent because products of the $f_i$ are. See the updated proof in the paper for the details.
\item[(25)]
It is not an ordering, only a partial ordering (which we have now stated).
\item[(27)]
Good point, we should have done a reduction instead of claiming it was analogous.
\item[(28)]
Wouldn't there be something like $\binom{m + c_1- 1}{c_1}$ elements? For example, if
we take $\alpha = 1$, and $H = V(x_0)$ we will have generators $x_1/x_0, \ldots, x_m/x_0$.
In any case, we decided to remove the paragraph this comment
was referring to and the claim about minimality
of the presentation from the theorem.
\end{enumerate}
\end{document}

### 2017-11-25 surfaces-write-up/MRL/MRL-surfaces-responses-to-referee-3.tex

\def\listtodoname{List of Todos}
\def\listoftodos{\@starttoc{tdo}\listtodoname}

\title{Summary of Changes in response to Referee's Report}
\author{Aaron Landesman, Peter Ruhm, and Robin Zhang}

\begin{document}

\maketitle
\section{Changes Made}
The authors would like to express their gratitude for
the numerous helpful suggestions given by the referee.
We applied nearly all such suggestions and give here
some brief comments on how we applied certain suggested
changes.
\begin{enumerate}
	\item[(2)] 
		We have added in a detailed example 
		(Example~\autoref{example:effective-example})
		explaining how to apply our method of proof to actually work out
		the generators and relations for a section ring
		associated to an effective divisor on $\bp^m$.

	\item[(10)]
		Response to the concern about base change along inseparable
		extensions:
		We added Remark~\autoref{rem:algebraically-closed}
		explaining why purely inseparable extensions pose
		no additional issue. The point is that cohomology commutes with
		flat base change, and extensions of fields
		are always flat, so the dimensions of the graded pieces of the
		ring will be preserved. Therefore generators
		and relations are preserved under arbitrary base field
		extension. Of course, canonical divisors of root stacks may not
		behave well with respect to base change (as we learned from a
		version of Voight--Zureick-Brown's forthcoming book available
		on Voight's website)
		which is why inseparable extensions
		may cause problems for canonical rings.
		But, for the purposes of this article, we fix a divisor on a
		scheme over some base field, and the structure of that particular section
		ring is unchanged upon base change to the algebraic closure.

	\item[(14) \& (18)]
		We explained in both the projective space case and Hirzebruch case
		that this assumption was nonessential, adding
		Corollary~\autoref{cor:proj-generators-relations} for the projective case
		and Remark~\autoref{rem:hirz-generators-relations-hypothesis} for the
		Hirzebruch case. You are correct that we could have easily changed
		coordinates and reordered the $f_i$ so that the first $m+1$
		many $f_i$'s were $x_0, \ldots, x_m$, but
		we may as well state it with arbitrary independent polynomials so
		as not to require such a change of basis.

	\item[(22)]
		We moved the notation to its own section after the introduction,
		as most of the notation was not necessary to state
		our main theorems.

	\item[(24)]
		Indeed, the proof of \autoref{prop:pm-span-and-basis}	was poorly
		phrased, and we have tried to clarify this now.
		The point is that we are only varying $f_0, \ldots, f_m,$ so
		the resulting monomials in the $f_i$ will be independent because
		products of the $f_i$ are. See the updated proof in the paper for the details.

	\item[(25)]
		It is not an ordering, only a partial ordering (which we have now stated).

	\item[(27)]
		Good point, we should have done a reduction instead of claiming it was analogous.
		This is now included at the beginning of the proof of the theorem
		(now numbered \autoref{thm:proj-one-point}).

	\item[(28)]
		Wouldn't there be something like $\binom{m + c_1- 1}{c_1}$ many
		elements? For example, if	we take $\alpha = 1$, and $H = V(x_0)$
		we will have generators $x_1/x_0, \ldots, x_m/x_0$.
		In any case, we decided to remove the paragraph mentioned by
		this comment as well as the claim about minimality of the
		presentation given in the theorem.
\end{enumerate}
\end{document}

### 2017-11-25 surfaces-write-up/surface-paper.tex
surfaces: revisions on Aaron's edits + finishing touches

Then the section ring $R_D$
is generated in degrees at most $\max_{0 \leq i \leq n}{k_i}$ with
relations generated in degrees at most $2 \max_{0 \leq i \leq n}{k_i}$.
\end{restatable}

\begin{restatable}{thm}{restateproj}
\label{thm:proj-generators-relations}
Let $D = \sum_{i = 0}^n \alpha_i D_i \in \di \bp^m \otimes_\bz \bq$
with $\alpha_i = \frac{c_i}{k_i}\in \bq$ in reduced form. Write
$\ell_i := \lcm_{0\le j \le n, j \ne i}(k_j)$ and $a_i := \deg D_i$. Let $\bp^m
\cong \proj \bk[x_0, \ldots, x_m]$ and let
$f_i \in \bk[x_0, \ldots, x_m]$ such that $D_i = V(f_i)$. 
Suppose that $\{f_0, \ldots, f_n\}$ contains as a subset forming a basis
for $H^0(\bp^m, \sco_{\bp^m}(1))$, i.e., $m+1$ independent polynomials with corresponding $a_i = 1$,
{\rm(}where we allow for 
$\alpha_j = 0$ in the expansion of $D${\rm)}.

Then $R_D$ is generated in degrees at most 
$\omega := \sum_{i = 0}^n \ell_i a_i$
with relations generated in degrees at most
\[
	\max \left(2 \omega, \frac{\max_{0 \le i \le n}
	(	\bida_i)}{\deg(D)} + \omega \right).
\]
\end{restatable}

@@

Then $R_D$ is generated in degrees at most 
$\omega' := \ell(m+1) + \sum_{i = 0}^n \ell_i a_i$
with relations generated in degrees at most
\[
	\max \left(2 \omega', \frac{\max_{0 \le i \le n}
	(	\bida_i)}{\deg(D)} + \omega' \right).
\]
\end{cor}

@@

\begin{thm}
\label{thm:hirz-generators-relations}
Let $D = \sum_{i = 1}^n \alpha_i D_i \in \di \hirz_m \otimes_\bz \bq$ where $\alpha_i = \frac{c_i}{k_i} \in \bq$ is written in reduced form.  
Let each $D_i = V(f_i)$, where $f_i \in \sco(a_i, b_i)$.
Let $u, v, z, w$ be the coordinates for the Hirzebruch surface
$\hirz_m$, and suppose that $\{f_1, \ldots, f_n\}$
contains 
bases for $\sco_{F_m}(1,0)$ and $\sco_{F_m}(0,1)$,
i.e., two independent linear polynomials in $u, v$ and two independent linear polynomials in
$w, x$ {\rm(}where we allow for $\alpha_j = 0$ in the expansion of $D${\rm)}.

Then $R_D$ is generated in degrees at most
\[
	\rho' := \lcm_{1\le i\le n}(k_i)\cdot \left(\sum_{1\le i\le j\le n} \bida_i \bidb_i\right) 
\]
\noindent
with relations generated in degrees at most 
$2 \rho'$.
%\[
%	2 \cdot \lcm_{1\le i\le n}(k_i) \cdot \left( \sum_{1\le i\le j\le n} \bida_i \bidb_i\right) 
%\]
\end{thm}
\begin{rem}
Theorem~\ref{thm:hirz-generators-relations} is restated with a more precise
bound in Theorem~\ref{thm:hirz-generators-relations-full}.
\end{rem}
\begin{rem}
	\label{rem:hirz-generators-relations-hypothesis}
	As in Theorem~\ref{thm:proj-generators-relations}, the condition that $D$ contain independent polynomials
	in $u,v$ and $w,z$ is easily removed by replacing $D$ with $D' :=D +0 \cdot u + 0 \cdot v + 0 \cdot w + 0 \cdot z$,
	and computing the resulting bound for $D'$.
	One can either take the simpler bound (but slightly weaker) bound from Theorem~\ref{thm:hirz-generators-relations}
	or the more precise bound given in Theorem~\ref{thm:hirz-generators-relations-full}.
\end{rem}

@@

In this section, we now collect various notation, used throughout the paper.
Throughout, for ease of notation,
we work over a fixed algebraically closed field $\bk$.
Our results hold equally well over arbitrary fields, as we now explain.

\begin{rem}
	\label{rem:algebraically-closed}
	Because cohomology commutes with flat base change, in particular with field extensions,
the dimensions of the graded pieces of the section ring will be preserved under base change from $\bk$ to $\overline \bk$. Therefore generators
and relations are preserved under arbitrary base field extension, and hence their minimal degrees are preserve.
Because of this, there is not harm in assuming $\bk =\overline \bk$. The bounds we give below hold equally well
over arbitrary fields, but for simplicity we state and prove them over algebraically closed fields.

We note that if $L/\bk$ is an inseparable extension,
and $X$ is a scheme over $\bk$ than the base change of the canonical divisor
$(K_X)_L$ may be different then the canonical divisor of the base change
$K_{X_L}$.
Therefore, the canonical ring may not be preserved under base change
along inseparable extensions.
But, for the purposes of this article, we fix a divisor on a scheme over some base field, and the structure of that particular section
ring is unchanged upon base change to the algebraic closure.
\end{rem}

We next define some notation used throughout the paper.
We use $m \in \bz_{\geq 0}$ to index the dimension of a given projective
space $\bp^m$ and the type of the Hirzebruch surface $\hirz_m$.
Let $D$ be a (Weil) $\bq$-divisor of the form
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \di X \otimes_\bz \bq.
\end{align*}\noindent
where $n \in \bz_{\geq 0}$ indexes the number of irreducible divisors in 
the above expansion of $D$, $\alpha_i \in \bq$, and $D_i \in \di X$ is an integral
codimension 1 closed subscheme of $X$. When it is convenient to do
so, we shall sometimes start the indexing at $0$, so that $i$ runs
from $0$ to $n$. In the case $X = \bp^m$, define the degree of $D$ by $\deg D :=
\sum_{i=1}^{n}\alpha_i \cdot \deg D_i$. The floor of a $\bq$-divisor $D$ is the divisor
$\lfloor D \rfloor := \sum_{i = 1}^{n}
\lfloor \alpha_i \rfloor D_i$. Let 
$R_D := \bigoplus_{d \geq 0} H^0(X, \lfloor dD \rfloor)$
denote the section ring associated to the $\bq$-divisor $D$. 
We often alternatively write
$R_D := \bigoplus_{d \geq 0} u^d H^0(X, \lfloor dD \rfloor)$, where $u$ is a dummy
variable to keep track of the degree.
We use $R_D$ as notation for $R(X,D)$ when $X$ is understood from context.
If $R$
is a graded ring, we denote the $d$th graded component of $R$ by $R_d$. If $r$ is a rational number, we let
$\fr(r) := r - \lfloor r
\rfloor$ denote the fractional part of $r$. 
If $D \in \di X \otimes_\bz \bq$
is an arbitrary divisor, we denote $h^0(X, D) := \dim_\bk H^0(X,D)$.

\section{Section Rings of Projective Space}
\label{sec:proj}
Let $\bk$ be a field and let $\bp^m$ denote $m$-dimensional
projective space over $\bk$. In this section, we bound the degrees
of generators and relations for divisors
on $\bp^m$, for all $m \geq 1$ in Theorem
~\ref{thm:proj-effective-intro}.
If $\deg D < 0$, the
section ring is concentrated in degree 0, and if
$\deg D = 0$, then the section ring has a single
generator. Therefore, for the remainder of this section,
we shall assume $\deg D > 0$.
The $\bp^1$ case restricts to the results of
\cite{dorney:canonical}. We also give an explicit description of the
generators of the section ring $R_D$ when $D$ is an effective divisor in Theorem ~\ref{thm:proj-generators-relations}.

@@

\noindent
be the convergents of the Hirzebruch-Jung continued fraction of
$\alpha$ (q.v. \cite[Section 2]{voight:cf} and \cite[Section 3]{hirzebruch:cf}). Then, the section ring
\[
	R_D := \bigoplus_{d \geq 0} u^d H^0(\bp^m, \lfloor dD \rfloor)
\]

@@

\begin{proof}
As a first step, we reduce to the case $D' = 0$.
Using that $D$ is effective, we can write
$R_D = R_{D'} + R_{\alpha H}$ (where the sum is nearly a direct sum, except 
the $(R_{D'})_d \cap (R_{\alpha H})_d$ is the one dimensional subspace generated by $u^d$).
Then, to give a presentation of $R_D$ over $R_D'$,
it suffices to give a presentation of $R_{\alpha H}$:
generators of $R_{\alpha H}$ map to generators of $R_D$ over $R_{D'}$ under the inclusion $\iota: R_{\alpha H} \rightarrow R_D$
and relations map to a full set of relations for $R_D$ over $R_{D'}$.
%is equivalent to giving a presentation
%For simplicity, we give a proof for when $D' = 0$, but the general case
%can be analogously proved by replacing the orders of zeros and poles
%by the order of zeros and poles relative to $D'$. That is,
%one can work with rational sections of $\mathscr O_{\bp^m}(D')$
%instead of rational sections of $\mathscr O_{\bp^m}$,
%and the following proof goes through otherwise unchanged.

Hence, for the remainder of the proof, we assume $D' = 0$.
%We will first check that $F^{\vec v}_i$ determine a system of minimal
%generators of $R_D$. We know $\left\{F_{1}^{\vec{v}} \colon \deg \vec{v} =
%\left\lfloor
%\frac{p}{q} \right\rfloor \right\}$ are minimal generators in
%degree $1$. Now let $i > 1$. Suppose $F_i^{\vec{v}}$ were not
%minimal for some $\vec{v} \in \mss_i$. Then $F_i^{\vec{v}} =
%F_{s}^{\vec{y}} F_{t}^{\vec{z}}$ for some $s, t \in \bn$ and
%$\vec{y}, \vec{z} \in \bz_{ \geq 0}^{m + 1}$. In particular, we see
%that 
%\begin{align*}
%	&d_i = s + t \\
%	&c_i = \deg \vec{v} = \deg \vec{y} + \deg \vec{z}
%\end{align*}
%
%\noindent
%so $\frac{c_i}{d_i}$ is the mediant of the rational numbers
%$\frac{\deg \vec{y}}{s}, \frac{\deg \vec{z}}{t}$ which both
%do not exceed $\alpha$. One of $\frac{\deg \vec{y}}{s},
%\frac{\deg \vec{z}}{t}$ must be at least $\frac{c_i}{d_i}$
%and $s$ and $t$ are both less than $d_i$, which contradicts the assumption that
%$\frac{c_i}{d_i}$ is a convergent of the Hirzebruch-Jung continued
%fraction of $\alpha$. Therefore, none of the $F_i^{\vec v}$ are redundant.
%

@@

The main aim of this section is to prove Theorem
~\ref{thm:proj-generators-relations} bounding the number of generators and
relations of $R_D$. 
Having justified the necessity of adding ghost divisors, we proceed to bound the
number of generators and relations of arbitrary $\bq$-divisors in 
projective space. 
In PRoposition~\ref{prop:pm-span-and-basis}, we record a general proposition describing
a basis for $H^0(\bp^m, dD)$.
We bound the generators in Lemma ~\ref{lem:proj-generators}, and we use Lemmas ~\ref{lem:composite-map}, ~\ref{lem:bound-ker-chi}, and ~\ref{lem:proj-relations-psi} to bound the 
degree of relations in the proof of Theorem ~\ref{thm:proj-generators-relations}. Proposition ~\ref{prop:cone-generation}, Lemma ~\ref{lem:composite-map}, and Lemma ~\ref{lem:bound-ker-chi} are quite 
general and will also be used in Section ~\ref{sec:hirz} to bound the 
degree of generators and relations on Hirzebruch surfaces.
However, before moving on to the proof of ~\ref{thm:proj-generators-relations}, we justify the importance of assumption
~\eqref{eqn:pm-basis-assumption} with several illustrative examples.

@@

We now develop the tools to bound the degrees of generators for $R_D$ with $D$ a divisor on $\bp^m$.

@@

We now move on to bounding the relations of a section ring $R_D$ for $D$ a divisor on $\bp^m$.
In Lemma ~\ref{lem:composite-map} we show that the degree of the
generators for $\ker \phi$, which is the same as the degree of relations
of $R_D$, is bounded by the maximum of the degree of generators for
$\ker \chi$ and for $\ker \psi$.
In Lemma ~\ref{lem:bound-ker-chi}, we bound the degree of generation of
$\ker \chi $, and we bound the degree of generation of $\ker \psi$ in Lemma ~\ref{lem:proj-relations-psi}

@@

\section{Section Rings of Hirzebruch surfaces}
\label{sec:hirz}
\invisiblesubsection
The aim of this section is to prove Theorem
~\ref{thm:hirz-generators-relations}, bounding the degree of
generators and relations of the section ring of a $\bq$-divisor
on any Hirzebruch surface.

@@

Furthermore, similarly to the case of projective space, 
we will restrict to the case that $D$ is a divisor
for which both of its bi-degrees are positive. We now justify this
restriction. If either of the bi-degrees of $D$ are negative,
then the section ring is concentrated in degree 0. If one of the
bi-degrees is 0, say the first one is 0, then $R_D$ is isomorphic
to $R_{D'}$, where $D' \in \di \bp^1 \otimes_\bz \bq$, 
where $D'$ can be written as a sum of divisors whose degrees
are multiples of the second bi-degree of $D$. Since the case of
$\bp^1$ has already been analyzed in \cite{dorney:canonical}, we
are justified in assuming that both bi-degrees of $D$ are positive.

>>>>

Then the section ring $R(\bp^m, D)$
is generated in degrees at most $\max_{0 \leq i \leq n}{k_i}$ with
relations generated in degrees at most $2 \max_{0 \leq i \leq n}{k_i}$.
\end{restatable}

\begin{restatable}{thm}{restateproj}
\label{thm:proj-generators-relations}
Let $D = \sum_{i = 0}^n \alpha_i D_i \in \di \bp^m \otimes_\bz \bq$
with $\alpha_i = \frac{c_i}{k_i}\in \bq$ in reduced form. Write
$\ell_i := \lcm_{0\le j \le n, j \ne i}(k_j)$ and $a_i := \deg D_i$. Let $\bp^m
\cong \proj \bk[x_0, \ldots, x_m]$ and let
$f_i \in \bk[x_0, \ldots, x_m]$ such that $D_i = V(f_i)$. 
Suppose that $\{f_0, \ldots, f_n\}$ contains a basis
for $H^0(\bp^m, \sco_{\bp^m}(1))$, {\rm(}i.e. $m+1$ independent polynomials with
corresponding $a_i = 1${\rm)}.

Then $R(\bp^m, D)$ is generated in degrees at most 
$\omega := \sum_{i = 0}^n \ell_i a_i$
with relations generated in degrees at most
\[
	\max \left(2 \omega, \frac{\max_{0 \le i \le n}
	(	\bida_i)}{\deg(D)} + \omega \right).
\]
\end{restatable}

@@

Then $R(\bp^m, D)$ is generated in degrees at most 
$\omega' := (m+1)\ell + \sum_{i = 0}^n \ell_i a_i$
with relations generated in degrees at most
\[
	\max \left(2 \omega', \frac{\max_{0 \le i \le n}
	(	\bida_i)}{\deg(D)} + \omega' \right).
\]
\end{cor}

@@

\begin{thm}
\label{thm:hirz-generators-relations}
Let $D = \sum_{i = 1}^n \alpha_i D_i \in \di \hirz_m \otimes_\bz \bq$ where $\alpha_i = \frac{c_i}{k_i} \in \bq$ is written in reduced form.  
Let each $D_i = V(f_i)$, where $f_i \in \sco(a_i, b_i)$.
Let $u, v, z, w$ be the coordinates for the Hirzebruch surface
$\hirz_m$, and suppose that $\{f_1, \ldots, f_n\}$
contains bases for $\sco_{F_m}(1,0)$ and $\sco_{F_m}(0,1)$
{\rm(}i.e., two independent linear polynomials in $u, v$ and two independent linear polynomials in
$w, z$ {\rm)}.

Then $R(\hirz_m, D)$ is generated in degrees at most
\[
	\rho' := \lcm_{1\le i\le n}(k_i)\cdot \left(\sum_{1\le i\le j\le n} \bida_i \bidb_i\right) 
\]
\noindent
with relations generated in degrees at most 
$2 \rho'$.
%\[
%	2 \cdot \lcm_{1\le i\le n}(k_i) \cdot \left( \sum_{1\le i\le j\le n} \bida_i \bidb_i\right) 
%\]
\end{thm}

\begin{rem}
Theorem~\ref{thm:hirz-generators-relations} is restated with a more
precise bound in Theorem~\ref{thm:hirz-generators-relations-full}.
\end{rem}

\begin{rem}
	\label{rem:hirz-generators-relations-hypothesis}
	As in Theorem~\ref{thm:proj-generators-relations}, the condition
	of Theorem~\ref{thm:hirz-generators-relations} that $D$ contain
	independent polynomials in $u, v$ and $w, z$ is
	easily removed by replacing $D$ with $D' :=D +0 \cdot u + 0 \cdot
	v + 0 \cdot w + 0 \cdot z$,	and computing the resulting bound for
	$D'$ using either the simpler (but slightly weaker) bound
	from Theorem~\ref{thm:hirz-generators-relations} or the more
	precise bound given in Theorem~\ref{thm:hirz-generators-relations-full}.
\end{rem}

@@

In this section, we now collect various notation used throughout the paper.
Throughout,
we work over a fixed algebraically closed field $\bk$ for ease of notation,
but our results hold equally well over arbitrary fields.

\begin{rem}
\label{rem:algebraically-closed}
Because cohomology commutes with flat base change (in particular
with field extensions) the dimensions of the graded pieces of the
section ring will be preserved under base change from $\bk$ to
$\overline \bk$. Therefore generators and relations are preserved
under arbitrary base field extension, and so their minimal
degrees are preserved. Consequently, there is no harm in
assuming $\bk = \overline \bk$ for our proofs. The bounds we give
hold equally well over arbitrary fields.

Note that if $L/\bk$ is an inseparable extension
and $X$ is a scheme over $\bk$, then the base change of the
canonical divisor $(K_X)_L$ may be different than the canonical
divisor of the base change $K_{X_L}$.
Therefore, the canonical ring may not be preserved under base change
along inseparable extensions. Nonetheless, the results we give are
not affected because given a divisor on a scheme over some base field,
the structure of that particular section ring is unchanged upon
base change to the algebraic closure.
\end{rem}

Let $D$ be a (Weil) $\bq$-divisor on a rational surface $X$ of the form
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \di X \otimes_\bz \bq.
\end{align*}\noindent
where $n \in \bz_{\geq 0}$ indexes the number of irreducible divisors in 
the above expansion of $D$, $\alpha_i \in \bq$, and $D_i \in \di X$ is an integral
codimension 1 closed subscheme of $X$. When it is convenient to do
so, we shall sometimes start the indexing at $0$, so that $i$ runs
from $0$ to $n$. In the case $X = \bp^m$, define the degree of $D$ by $\deg D :=
\sum_{i=1}^{n}\alpha_i \cdot \deg D_i$. The floor of a $\bq$-divisor $D$ is the divisor
$\lfloor D \rfloor := \sum_{i = 1}^{n}
\lfloor \alpha_i \rfloor D_i$.

Let $R(X, D) := \bigoplus_{d \geq 0} H^0(X, \lfloor dD \rfloor)$
denote the section ring associated to the $\bq$-divisor $D$. 
We often alternatively write
$R(X, D) := \bigoplus_{d \geq 0} u^d H^0(X, \lfloor dD \rfloor)$, where $u$ is a dummy
variable to keep track of the degree.
When $X$ is understood from context, we use $R_D$ as notation for $R(X, D)$.

We use $m \in \bz_{\geq 0}$ to index the dimension of a given
projective space $\bp^m$ and the type of the Hirzebruch surface
$\hirz_m$. If $S$ is a graded ring, we denote the $d$th graded
component of $S$ by $S_d$. If $r$ is a rational number, we let
$\fr(r) := r - \lfloor r \rfloor$ denote the fractional part of $r$. 
If $D \in \di X \otimes_\bz \bq$ is an arbitrary divisor, we denote
$h^0(X, D) := \dim_\bk H^0(X,D)$.

\section{Section Rings of Projective Space}
\label{sec:proj}
Let $\bk$ be a field and let $\bp^m$ denote $m$-dimensional
projective space over $\bk$. In this section, we prove
Theorem~\ref{thm:proj-generators-relations}, which bounds the degrees
of generators and relations of the section ring of any $\bq$-divisor
on $X = \bp^m$ for all $m \geq 1$. We also prove
Theorem ~\ref{thm:proj-effective-intro} to give an explicit
description of the generators of the section ring $R_D$ when $D$ is
an effective divisor.

If $\deg D < 0$, the
section ring is concentrated in degree 0, and if
$\deg D = 0$, then the section ring has a single
generator. Therefore, for the remainder of this section,
we shall assume $\deg D > 0$.
Note that the $\bp^1$ case, in particular, restricts to the results
of \cite{dorney:canonical}.

@@

\noindent
be the convergents of the Hirzebruch-Jung continued fraction of
$\alpha$ {\rm(}q.v. \cite[Section 2]{voight:cf} and
\cite[Section 3]{hirzebruch:cf}{\rm)}. Then, the section ring
\[
	R_D := \bigoplus_{d \geq 0} u^d H^0(\bp^m, \lfloor dD \rfloor)
\]

@@

\begin{proof}
As a first step, we reduce to the case $D' = 0$.
Using that $D$ is effective, we can write
$R_D = R_{D'} + R_{\alpha H}$ (where the sum is nearly a direct sum, except 
the $(R_{D'})_d \cap (R_{\alpha H})_d$ is the one dimensional subspace generated by $u^d$).
Then, to give a presentation of $R_D$ over $R_D'$,
it suffices to give a presentation of $R_{\alpha H}$:
generators of $R_{\alpha H}$ map to generators of $R_D$ over $R_{D'}$ under the inclusion $\iota: R_{\alpha H} \rightarrow R_D$
and relations map to a full set of relations for $R_D$ over $R_{D'}$.
Hence, for the remainder of the proof, we can assume $D' = 0$.
%is equivalent to giving a presentation
%For simplicity, we give a proof for when $D' = 0$, but the general case
%can be analogously proved by replacing the orders of zeros and poles
%by the order of zeros and poles relative to $D'$. That is,
%one can work with rational sections of $\mathscr O_{\bp^m}(D')$
%instead of rational sections of $\mathscr O_{\bp^m}$,
%and the following proof goes through otherwise unchanged.

%We will first check that $F^{\vec v}_i$ determine a system of minimal
%generators of $R_D$. We know $\left\{F_{1}^{\vec{v}} \colon \deg \vec{v} =
%\left\lfloor
%\frac{p}{q} \right\rfloor \right\}$ are minimal generators in
%degree $1$. Now let $i > 1$. Suppose $F_i^{\vec{v}}$ were not
%minimal for some $\vec{v} \in \mss_i$. Then $F_i^{\vec{v}} =
%F_{s}^{\vec{y}} F_{t}^{\vec{z}}$ for some $s, t \in \bn$ and
%$\vec{y}, \vec{z} \in \bz_{ \geq 0}^{m + 1}$. In particular, we see
%that 
%\begin{align*}
%	&d_i = s + t \\
%	&c_i = \deg \vec{v} = \deg \vec{y} + \deg \vec{z}
%\end{align*}
%
%\noindent
%so $\frac{c_i}{d_i}$ is the mediant of the rational numbers
%$\frac{\deg \vec{y}}{s}, \frac{\deg \vec{z}}{t}$ which both
%do not exceed $\alpha$. One of $\frac{\deg \vec{y}}{s},
%\frac{\deg \vec{z}}{t}$ must be at least $\frac{c_i}{d_i}$
%and $s$ and $t$ are both less than $d_i$, which contradicts the assumption that
%$\frac{c_i}{d_i}$ is a convergent of the Hirzebruch-Jung continued
%fraction of $\alpha$. Therefore, none of the $F_i^{\vec v}$ are redundant.
%

@@

The main aim of this section is to prove Theorem
~\ref{thm:proj-generators-relations}.
Having justified the necessity of adding ghost divisors, we proceed to bound the
number of generators and relations of arbitrary $\bq$-divisors in 
projective space. 
In Proposition~\ref{prop:pm-span-and-basis}, we record a general proposition describing
a basis for $H^0(\bp^m, dD)$.
We bound the generators in Lemma ~\ref{lem:proj-generators}, and we use Lemmas ~\ref{lem:composite-map}, ~\ref{lem:bound-ker-chi}, and ~\ref{lem:proj-relations-psi} to bound the 
degree of relations in the proof of Theorem ~\ref{thm:proj-generators-relations}. Proposition ~\ref{prop:cone-generation}, Lemma ~\ref{lem:composite-map}, and Lemma ~\ref{lem:bound-ker-chi} are quite 
general and will also be used in Section ~\ref{sec:hirz} to bound the 
degree of generators and relations on Hirzebruch surfaces.
However, before moving on to the proof of ~\ref{thm:proj-generators-relations}, we justify the importance of assumption
~\eqref{eqn:pm-basis-assumption} with several illustrative examples.

@@

We now develop the tools to bound the degrees of generators for $R_D$
for $D$ an arbitrary divisor on $\bp^m$.

@@

We now move on to bounding the degree of relations of a section ring $R_D$ for $D$ a divisor on $\bp^m$.
In Lemma ~\ref{lem:composite-map} we show that the degree of the
generators for $\ker \phi$, which is the same as the degree of relations
of $R_D$, is bounded by the maximum of the degree of generators for
$\ker \chi$ and for $\ker \psi$.
In Lemma ~\ref{lem:bound-ker-chi}, we bound the degree of generation of
$\ker \chi $, and we bound the degree of generation of $\ker \psi$ in Lemma ~\ref{lem:proj-relations-psi}

@@

\section{Section Rings of Hirzebruch Surfaces}
\label{sec:hirz}
\invisiblesubsection
Let $\hirz_m$ denote the $m$-th Hirzebruch surface.
The aim of this section is to prove Theorem
~\ref{thm:hirz-generators-relations}, which bounds the degree of
generators and relations of the section ring of any $\bq$-divisor
on $X = \hirz_m$, for all $m \geq 0$.

@@

Furthermore, we will restrict to the case that $D$ is a divisor
for which both of its bi-degrees are positive. We now justify this
restriction. If either of the bi-degrees of $D$ are negative,
then the section ring is concentrated in degree 0. If one of the
bi-degrees is 0, say the first one is 0, then $R_D$ is isomorphic
to $R_{D'}$, where $D' \in \di \bp^1 \otimes_\bz \bq$, 
where $D'$ can be written as a sum of divisors whose degrees
are multiples of the second bi-degree of $D$. Since the case of
$\bp^1$ has already been analyzed in \cite{dorney:canonical}, we
are justified in assuming that both bi-degrees of $D$ are positive.
